---
fontsize:10
---
**Statistical Inference Programming Assignment Part 2**
Author: "Hamel Husain"

This is the report for part 2 (inference) of the Coursera statistical inference class.  The full code for this report can be viewed at: 

**https://github.com/hamelsmu/datasciencecoursera/tree/master/StatInference**

**1.  Load the ToothGrowth data and provide basic summary of data. **
```{r echo = FALSE}
library(datasets)
library(ggplot2)
data(ToothGrowth) 
str(ToothGrowth)
```

**2.  Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.**
I will first explore the data via density plot:
```{r echo = FALSE, fig.width=10, fig.height=5}
#Make Plot
g = ggplot(ToothGrowth, aes(len))
g + geom_density(aes(fill = supp), alpha = .5) + facet_grid(dose~.) + 
labs(x = 'Tooth Length', y = 'Density', 
title = 'The Effect of Vitamin C on Tooth Growth\nin Guinea Pigs by Dosage (.5, 1, 2) and\nMethod of Delivery (OJ, VC)', 
fill = 'Method of Delivery\nOJ:Orange Juice\nVC:Absorbic Acid')
```

Based on the the density plot, I have several hypothesis that I would like to test.  **(1)** For a medium dose (dose 1), Vitaman C adminstered through OJ is associated with longer tooth length than VC.  **(2)** Controlling for delivery method, higher doses lead to higher tooth length.  

**Test Hypothesis 1**
```{r}
OJ = ToothGrowth[ToothGrowth$dose ==1 & ToothGrowth$supp == 'OJ', ]$len
VC = ToothGrowth[ToothGrowth$dose ==1 & ToothGrowth$supp == 'VC', ]$len
t.test(OJ, VC)
```

**Conclusion**: Based on these results, I reject the null hypothesis that the mean difference between OJ and VC where dose = 1 is zero.  I feel confident in saying that OJ is associated with longer teeth length than VC for medium dosage.  I am assuming that the variances of these two groups are equal.

**Test Hypothesis 2**
```{r}
low_dose_OJ = ToothGrowth[ToothGrowth$dose ==.5 & ToothGrowth$supp == 'OJ', ]$len
high_dose_OJ = ToothGrowth[ToothGrowth$dose ==2 & ToothGrowth$supp == 'OJ', ]$len
low_dose_VC = ToothGrowth[ToothGrowth$dose ==.5 & ToothGrowth$supp == 'VC', ]$len
high_dose_VC = ToothGrowth[ToothGrowth$dose ==2 & ToothGrowth$supp == 'VC', ]$len
t.test(high_dose_OJ, low_dose_OJ)
t.test(high_dose_VC, low_dose_VC)
```

**Conclusion**: I reject the null hypothesis that high dosages have the same mean tooth length as low dosages, when controlling for delivery method.  Furthermore, I noticed that the effect of dosage is more prounounced when the delivery method is VC.  I again assumed the variance of the two groups were equal when performing the t-test.
