##Statistical Inference - Simulation
Author: Hamel Husain

This is the report for part 1 (simulation) of the Coursera statistical inference class.  The problem statement is as follows:

***The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Set lambda = 0.2 for all of the simulations. ***

First, I will collect 1,000 sample means of size n=40 from a simulated dataset that has exponential distribution with lambda = .2

```{r simulate_data}
set.seed(1)
library(data.table)
#define function to collect sample mean of n=40 from aforementioned distribution
sim_exp = function(x){return(x * mean(rexp(40, .2)))}

#Collect 1000 sample means of size n=40 with lambda = .2
sim = apply(data.table(rep.int(1, 1000)), 1, sim_exp)
```

#####1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
#####2. Show how variable it is and compare it to the theoretical variance of the distribution.  

These two questions can be answered by simply calculating the observed and theoretical mean and variance.  The theoretical variance is the square of the theoretical standard deviation, which is 1/lambda divided by the square root of the sample size n =40. The theoretical mean is 1/lambda.
```{r theoretical}
theoretical_variance = round(((1/.2) / sqrt(40))^2,3)
actual_variance = round(var(sim),3)
theoretical_mean = 1/.2  #theo mean = 1/lambda
actual_mean = round(mean(sim),3)
compare_var = rbind(theoretical_mean,actual_mean,theoretical_variance, actual_variance)
compare_var #Display the difference
```

As we can see, the theoretical variance and observed variance are very similar.  Furthermore, the theoretical mean and actual mean are also very similar.

#####3. Show that the distribution is approximately normal.
```{r normal_histogram, echo=FALSE}
#Generate Histogram
hist(sim, prob = T, main = 'Density Plot of 1,000 Sample Means\n Exponential Distribution n=40, lamda = .2',
     xlab = 'Sample Mean (n=40)',
     cex.main = .9,
     ylim = c(0,.6),
     breaks = 20
)

curve(dnorm(x, mean(sim), sd(sim)), 
           add = TRUE,
           col = 'darkblue', lwd = 2)

lines(density(sim), lty="dotted", col="red", lwd=3) 

#create legend
legend('topright', lty =c(1, 3), lwd = 3, col = c('blue', 'red'),
       cex = .75,
       legend = c('Normal Distribution', 'Observed Distribution')
)
```

As we can see from this graph, the observed density is very similar to the normal density, and thus it is fair to say that the observed data is approximately normal.

#####4. Evaluate the coverage of the confidence interval for 1/lambda: X ± 1.96 (s/√n).
```{r}
#calculate the confidence interval
ci = mean(sim) + (c(-1,1)* (1.96 * ((1/.2) / sqrt(40))))
#calculate the coverage
sum(between(sim, ci[1], ci[2])) / length(sim)
```

The coverage is approximately 94.9%.  This is expected as the data is normally distributed.

---